\chapter*{Glossary}
\begin{itemize}
\item \textbf{GSP}: Generalized Steiner Problem.
\item \textbf{GSP-NC}: GSP with Node-Connectivity Constraints.
\item \textbf{GSP-EC}: GSP with Edge-Connectivity Constraints.
\item \textbf{GSP-NCHR}: GSP with Node-Connectivity Constraints and Hostile Reliability.
\item \textbf{GSP-ECHR}: GSP with Edge-Connectivity Constraints and Hostile Reliability.
\item \textbf{SNDP}: Survivable Network Design Problem.
\item \textbf{GSNDP}: Generalized Survivable Network Design Problem.
\item \textbf{GNDP}: Generalized Network Design Problem.
\item \textbf{VCSNDP}: Vertex Connectivity Survivable Network Design Problem.
\item \textbf{SN-MSP}: Survivable Network with Minimal Steiner nodes Problem.
\item \textbf{SMT}: Steiner Minimal Tree.
\item \textbf{MCSP}: Minimal Connected Sub Graph Problem.
\item \textbf{MST}: Minimum Spanning Tree.
\item \textbf{TSP}: Travel Salesman Problem.
\item \textbf{BNDP}: Backbone Network Design Problem.
\item \textbf{KSP}: K Shortest Paths.
\item \textbf{WAN}: Wide Area Network.
\item \textbf{IP}: Internet Protocol.
\item \textbf{MPLS}: Multiprotocol Label Switching.
\item \textbf{ZDD}: Zero Decision Diagram.
\item \textbf{CMC}: Crude Monte Carlo.
\item \textbf{RVR}: Recursive Variance Reduction.
\item \textbf{RNN}: Random Neural Network.
\item \textbf{SMBS}: Stochastic Monotone Binary System.
\item \textbf{GRASP}: Greedy Randomized Adaptive Search Procedure.
\item \textbf{IoT}: Internet of Things.
\item \textbf{GA}: Genetic Algorithm.
\item \textbf{RL}: Relocation Heuristics.
\item \textbf{VNS}: Variable Neighborhood Search.
\item \textbf{ILS}: Iterated Local Search.
\item \textbf{TS}: Tabu Search.
\item \textbf{VND}: Variable Neighborhood Descent.
\item \textbf{VNDS}: Variable Neighborhood Decomposition Search.
\item \textbf{BVNS}: Biased Variable Neighborhood Search.
\item \textbf{PVNS}: Parallel Variable Neighborhood Search.
\end{itemize} 